---
title: "Preliminary Summary of Kenya Data"
author: "Michael Flynn"
---


```{r setup}
#| echo: false
#| include: false
#| message: false
#| warning: false
#| error: false

library(tidyverse)
library(psych)
library(purrr)
library(targets)
library(kableExtra)

#tar_source("../R/")
tar_load(survey_clean)
```

# Notes

1. Not a lot of variation in the amount of influence questions. Most responses heavily concentrated on "some" and "a lot". More variation on the quality of influence.  

2. Interestingly high number of people saying they don't know when asked if China has a military presence in Kenya. Notably, some of the people saying "don't know" here seem to be providing answers other than "don't know" when asked about the quality of influence.

3. Way higher rate of contact with Chinese citizens than US. 

4. Income is highly skewed to the lowest income grouping. Roughly 65% of respondents fall into the lowest income quantile. 

5. Age is fairly even-ish across the three groupings, but that 35+ category is catching a lot of variation. We also have the specific age variable so we can generate new groupings if need be.

6. Respondents overwhelmingly from Mombasa. 


# Tables and Figures

```{r overview-tables}
#| echo: false
#| warning: false

varlist <- colnames(survey_clean)

outlist <- list()

for (i in varlist){
  
    kableExtra::kbl(print(table(survey_clean[, i])),
                    caption = paste0("Variable Name: ", i),
                    format = "html")
}

outlist <- map(
  .x = seq_along(varlist),
  .f = ~ ggplot(data = survey_clean, aes(x = survey_clean[[.x]])) +
    geom_bar() +
    theme(axis.text.x.bottom = element_text(angle = 45, vjust = 1, hjust = 1)) +
    labs(title = glue::glue("{varlist[[.x]]}"))
)

outlist

```
